---
title: "More info on ggplots"
date: "2016-08-27"
slug: "more-info-on-ggplots"
categories: ["R", "ggplot2"]
tags: ["ggplot2", "ggplot_build()"]
---

I have been a big fan of [Hadley Wickham's](http://hadley.nz) [ggplot](http://ggplot2.org) package for a while. This last week I was building a [shiny](http://shiny.rstudio.com) app and wanted to find out exactly which colours were being used in the plot. After a bit of googling I ended up at this [stackoverflow](http://stackoverflow.com/questions/11774262/how-to-extract-the-fill-colours-from-a-ggplot-object) post, which led me to the `ggplot_build` function which I found seriously useful.

A little demo:


```{r}
library(ggplot2)

# create a sample plot
p = ggplot(mtcars, aes(x = hp, y = mpg, colour = factor(cyl), shape = factor(am))) +
  geom_point(size = 3) +
  labs(colour = "No. Cylinders", shape = "Auto/Manual",
       x = "Horsepower", y = "Miles per Gallon",
       title = "Scatterplot of Horsepower vs Miles per Gallon")
# and display it
p
```

One way we can look at what info `ggplot` is using to make the plot is with the `ggplot_build()` function. This outputs a list of, amongst other things, a data frame of all the aesthetics. 


```{r}
# we can use the ggplot_build function to have a look at exactly what information ggplot is using
# to construct the plot
ggplot_build(p)$data
```


I found this very useful to double check that all the things I was expecting to happen, were actually happening and in to solve my particular problem of finding out exactly which colours I was using. For more detail on all the information `ggplot_build` can provide have a look at the [docs](http://docs.ggplot2.org/current/ggplot_build.html).
