---
title: "Dealing with missing data in ggplot2 barplots"
date: "2017-09-20"
slug: "missing-data-ggplot2-barplots"
categories: ["R"]
tags: ["ggplot2", "geom_bar()"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = NA, 
                      message = FALSE, warning = FALSE)
```

I'm wrting this down because it's a problem I've tried to deal with twice in the last couple of weeks and I had to look the solution up both times. The second time, rather than just fix it I dug a little deeper to find out why something I thought would work didn't. 

Any way, the problem:

```{r the-problem}
library(ggplot2)
fake_data <- data.frame(var1 = c("a", "a", "a", "b", "b", "b", "c"), 
                        var2 = c("x", "x", "y", "x", "y", "y", "x"))
fake_data 
ggplot(fake_data, aes(x = var1, fill = var2)) +
  geom_bar(position = "dodge")
```

Now that all works fine but I don't want the bar for `c` to be full width, just half width as the others are, so the missing data (or count of zero is clearer). 

This same problem has been dealt with in two stackoverflow posts^[https://stackoverflow.com/questions/11020437/consistent-width-for-geom-bar-in-the-event-of-missing-data] ^[https://stackoverflow.com/questions/10326729/dont-drop-zero-count-dodged-barplot] and Hadley Wickham (the package author) commented on the design of the `position_dodge()` in this [issue](https://github.com/tidyverse/ggplot2/issues/1776). The basic idea is to stop `ggplot` from dropping the missing bar (or bar with count 0). 

```{r}
ggplot(fake_data, aes(x = var1, fill = var2)) +
  geom_bar(position = "dodge") +
  scale_x_discrete(drop = FALSE) +
  scale_fill_discrete(drop = FALSE)
```

As you can see this makes no difference. The reason is because all the solutions count the data before plotting and use `stat = "identity"` and I wondered why that was necessary, why wouldn't it work without it. To get the answer I had a look at the plot in more detail.

```{r}
p = ggplot(fake_data, aes(x = var1, fill = var2)) +
  geom_bar(position = "dodge")
ggplot_build(p)$data
```

The data only has 5 rows and so there isn't any data to stop `ggplot` from dropping. It is necessary to count the data first. 

```{r}
# starting again
library(tidyverse)
fake_data <- data.frame(var1 = c("a", "a", "a", "b", "b", "b", "c"), 
                        var2 = c("x", "x", "y", "x", "y", "y", "x"))
# count the data
count_fake = count(fake_data, var1, var2)
# fill in the missing combination
count_fake = complete(count_fake, var1, var2) 
# plot the data
ggplot(count_fake, aes(x = var1, y = n, fill = var2)) +
  geom_bar(stat = "identity", position = "dodge")
```

**Note:** In the development version of `ggplot2` there is a much easier solution. First we need to install the development version (presuming you have already installed `devtools`.)

```{r, eval = FALSE}
devtools::install_github("tidyverse/ggplot2")
```

Then we have access to an additional argument within `position_dodge()`. 

```{r}
library(ggplot2)
fake_data <- data.frame(var1 = c("a", "a", "a", "b", "b", "b", "c"), 
                        var2 = c("x", "x", "y", "x", "y", "y", "x"))
ggplot(fake_data, aes(x = var1, fill = var2)) +
  geom_bar(position = position_dodge(preserve = 'single'))
```

