---
title: Postcode Polygons on a Leaflet Map
author: ''
date: '2018-05-30'
slug: postcode-polygons-on-a-leaflet-map
categories:
  - R
tags:
  - leaflet
description: ''
highlight: true
bibliography: "../../../blogdown_datasets/aj2duncan.bib"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, comment = NA, 
                      warning = FALSE, message = FALSE)
library(rgdal)
library(leaflet)
postcodes = readOGR("../../../blogdown_datasets/postcode_polygons.gpkg", 
                    layer = "postcode_district")
```

I have been using [Leaflet](https://rstudio.github.io/leaflet/) for a little while now but always as part of a __shiny__ app. However, as [Mike Spencer](http://mikerspencer.com/) went to all the trouble of producing polygons for [GB postcode areas and districts](https://scottishsnow.wordpress.com/2018/03/28/gb-postcode-polygons-open-data/) I thought this was a good opportunity for a blog post [@Spencer2018]. 

## Sorting the Data

First we'll load the packages we need.

```{r, eval = FALSE}
library(rgdal)
library(leaflet)
```

Next we read in the data. The file has two layers of polygons and we need to specify which one we want. I have chosen to work with the postcode district which is the first half of our postcodes. 

```{r, eval = FALSE}
postcodes = readOGR("postcode_polygons.gpkg", layer = "postcode_district")
```

For leaflet to be able to plot these they need to be in terms of *latitude* and *longitude* rather than *Easting* and *Northing*. We can check the current projection:

```{r}
proj4string(postcodes)
```

and convert them to *latitude* and *longitude*.

```{r}
postcodes = spTransform(postcodes, CRS("+proj=longlat +datum=WGS84"))
proj4string(postcodes)
```

For a lot more detail on different coordinate systems [this pdf](https://www.bnhs.co.uk/focuson/grabagridref/html/OSGB.pdf) seems like a reasonable place to start. 

## Plotting the Ploygons

To simplify plotting things (and reduce the amount of data needed for the leaflet map) we can select just those postcode districts starting **IV**. 

```{r}
postcodes = subset(postcodes, grepl("IV", postcodes$pc_district))
```

Now we can construct a map

```{r}
leaflet(postcodes, width = "100%") %>% 
  setView(lng = -3.64, lat = 58, zoom = 7) %>%
  addTiles() %>% 
  addPolygons()
```

To get a little more info we can add the postcode district as a popup. Changing the highlight options and adding a little custom css to the popup improves how everything looks. 

```{r, fig.align = "center"}
leaflet(postcodes, width = "100%") %>% 
  setView(lng = -3.64, lat = 58, zoom = 7) %>%
  addTiles() %>% 
  addPolygons(fillOpacity = 0.2, popup = ~(pc_district), weight = 2,
              highlightOptions = highlightOptions(color = "#03F", weight = 5, 
                                                  bringToFront = TRUE))
```

For info the custom popup css is shown below.

```{css}
.leaflet-popup-content-wrapper {
  background:rgba(0,51,255,0.5);
  color:rgba(255,255,255,0.7);
  font-size:14px;
  line-height:20px;
  border-radius: 0px;  
}
.leaflet-popup-tip {
  background:rgba(0,51,255,0.5);
  border-left:2px solid transparent;
  border-right:2px solid transparent;
  border-top:2px solid #03F; 
}
```

This was a good bit easier than I was expecting it to be ^[Not sure why I thought it was going to be difficult but there you go...] but mapping anything just reminds me of a [particular West Wing episode](https://www.youtube.com/watch?v=OH1bZ0F3zVU).

# References
