---
title: The Value of a Minimal Reproducible Example
author: ''
date: '2018-05-16'
slug: the-value-of-a-minimal-reproducible-example
categories:
  - R
tags: []
description: ''
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, comment = NA, 
                      warning = FALSE, message = FALSE)
```

I'm writing this down in the hope that it will remind me in the future. 

Although I have used a [minimal reproducible example](https://gist.github.com/hadley/270442) before, I found one particularly useful the other day. 

### The Problem

I had developed a [shiny](https://shiny.aj2duncan.com) app and part of this was a [leaflet](https://rstudio.github.io/leaflet/) map where users are able to click the map to add a marker, these are later captured. However, the markers always appeared below and to the right of where the point clicked. 

Checking the usual sources of helpful advice (Google, Stackoverflow) yielded nothingm, so I started from scratch.

### The Solution

Really quite quickly I found the answer. Due to running the app on a shiny server, behind ssl I was using a custom declaration for the markers to make sure they were also using ssl. 

```{r, eval = FALSE}
markers <- makeIcon(
  iconUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/0.7.7/images/marker-icon.png",
  shadowUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/0.7.7/images/marker-shadow.png"
)
```

What I had neglected to do was to take into account that by default [leaflet](https://rstudio.github.io/leaflet/) puts the marker image with the top left of it at the actual latitude and longitude you want marked. Setting the anchor points sorted everything out.

```{r, eval = FALSE}
markers <- makeIcon(
  iconUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/0.7.7/images/marker-icon.png",
  iconAnchorX = 12, iconAnchorY = 41,
  shadowUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/0.7.7/images/marker-shadow.png",
  shadowAnchorX = 12, shadowAnchorY = 41
)
```

I don't know how long this would have taken if I hadn't built it up piece by piece - keeping it as minimal as possible. Hopefully I'll remember how useful the process was next time around. 

Full code for the minimal shiny app is here:

```{r, eval = FALSE}
library(shiny)
library(leaflet)

ui <- fluidPage(
  fluidRow(
    column(12, verbatimTextOutput("click"))
  ),
  fluidRow(
    column(12, leafletOutput("map"))
  )
)


server <- function(input, output) {
   
  output$map <- renderLeaflet({
    leaflet("map") %>%
      setView(lng = -3.64, lat = 57.36, zoom = 8) %>%
      addTiles() 
  })
  
  observeEvent(input$map_click, { 
    markers <- makeIcon(
      iconUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/0.7.7/images/marker-icon.png",
      iconAnchorX = 12, iconAnchorY = 41,
      shadowUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/0.7.7/images/marker-shadow.png",
      shadowAnchorX = 12, shadowAnchorY = 41
    )
    
    leafletProxy("map") %>%
      addMarkers(lng = input$map_click$lng, lat = input$map_click$lat,
                 icon = markers)
  })
  
  observeEvent(input$map_click, {
    output$click <- renderPrint(input$map_click)
  })
}

# Run the application 
shinyApp(ui = ui, server = server)
```

