---
title: Tibbles, Data frames and Vectors
author: ''
date: '2019-06-18'
slug: tibbles-dataframes-and-vectors
categories:
  - R
tags:
  - tibble
description: 'Learning something new about tibbles and data frames'
highlight: true
---


```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, comment = NA, 
                      warning = FALSE, message = FALSE)
```

I learned something new today and the reason I write this is that I'm somewhat confused as to how I didn't know it already. 

# Data Frames

Let us start by creating a data frame. 

```{r}
temp = data.frame(x = c(1:3), y = c(10:12))
temp
```

We can select rows

```{r}
temp[1,]
```

and columns

```{r}
temp$x
temp[,2]
```

What I hadn't realised before is that there is another way to select columns, using `[[` notation that I'm used to using with lists. 

```{r}
temp[[1]]
```

Let's see if there is any difference in what is returned with these variations.

```{r}
class(temp[1,])
class(temp[,1])
class(temp$x)
class(temp[[1]])
```

So selecting a row returns a *data frame*, whereas selecting a column in any way seems to return a *vector*. 

# Tibbles

Let's do the same things with a tibble

```{r}
library(tibble)
temp2 = tibble(x = c(1:3), y = c(10:12))
temp2
```

We can select rows

```{r}
temp2[1,]
```

and columns

```{r}
temp2$x
temp2[,2]
temp2[[1]]
```

You've probably noticed a difference already but if we now look at what is returned we see some differences to the behaviour of `data.frame()`. 
 
```{r}
class(temp2[1,])
class(temp2[,1])
class(temp2$x)
class(temp2[[1]])
```

There is now consistency using `[,]` notation - both `temp2[1,]` and `temp2[,1]` return a *tibble*. Using either `$` or `[[]]` returns a vector as it did for a *data.frame*. 

# Why did this make a difference?

What happened in my case was that I'd written a function that had a *data frame* as one of the arguments. In this function I'd used the `$` notation to select a column and then treated this as a *vector*. Not a problem. 

I then wrote a similar function and used the `[,]` notation to select a column and then treated it as a *vector*. This wasn't a problem when I gave it a *data frame* but then I gave it a **tibble**. This actually threw an error for me but this sort of thing could result in just the wrong value being returned. A quick demo of this could be using the `length()` function. 

```{r}
# data frame
length(temp[,1])
# tibble
length(temp2[,1])
```

You may be wondering what sort of error I got, well as an example lets do something slightly more meaningful. 

```{r}
temp = data.frame(x = rnorm(100))
temp2 = tibble(x = rnorm(100, 2, 0.5))
```

Then something like this will work fine

```{r}
t.test(temp$x, temp2$x)
```

but this throws an error.

```{r, error = TRUE}
t.test(temp[,1], temp2[,1])
```

Now I know about `[[]]` I will definitely being using it with *tibbles*. 