---
title: Leaflet Draw, Exporting Polygons and the Holepunch Package
author: ''
date: '2019-11-06'
slug: leaflet-draw-exporting-shapefiles-and-the-holepunch-package
categories:
  - R
  - Shiny
tags:
  - leaflet
highlight: true
description: 'Using a Shiny App to Draw a POlygon and Export it'
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, comment = NA, 
                      warning = FALSE, message = FALSE)
```


A little over a month ago I attended [FOSS4GUK 2019](https://uk.osgeo.org/foss4guk2019/){target='_blank'}. It was a great conference and there were a few presentations (like mine) on [Leaflet](https://rstudio.github.io/leaflet/){target="_blank"}. One conversation I had was about using [Leaflet](https://rstudio.github.io/leaflet/){target="_blank"} to gather data and the question was whether it could be used draw and export polygons/polylines. I thought I would see if it could be. I have previously used a shiny app to record points submitted by users but not to record polygons.

## Leaflet

We can initialize a map quite easily.

```{r}
library(leaflet)
leaflet() %>%
  setView(lng = -4, lat = 58, zoom = 7) %>%
  addTiles()
```

We can also use the [Leaflet Extras](https://bhaskarvk.github.io/leaflet.extras/){target="_blank"} package to add [Leaflet Draw](https://github.com/Leaflet/Leaflet.draw){target="_blank"} to it. 

```{r}
library(leaflet)
library(leaflet.extras)
leaflet() %>%
  setView(lng = -4, lat = 58, zoom = 7) %>%
  addTiles() %>%
  addDrawToolbar()
```

This now gives users the possibility of drawing polygons. If we embed one of these in a shiny app then we are able to discover when the user has completed the polygon by using `observeEvent(input$map_draw_new_feature, {...})`. Having done this we can access the geometry of the polygon using `input$map_draw_new_feature$geometry$coordinates[[1]]`, convert to a shapefile and allow the user to download it.

## Holepunch

Having got a shiny app that lets uers draw and download a polygon I was trying to figure how to publish it. I could have put it on my [shinyserver](https://shiny.aj2duncan.com) but then I read about [binder](https://mybinder.org/){target="_blank"} and the [holepunch package](https://github.com/karthik/holepunch){target="_blank"}. Binder is a resource that lets you turn almost any Github repository into a open version of RStudio or Jupyter notebooks. The holepunch package helps you turn your `R` based Github repo into an RStudio instance, including any dependencies. The documentation is available [here](https://karthik.github.io/holepunch/){target="_blank"} but the process is pretty straightforward. At some point I may document the whole thing but for the moment I will leave it to the documentation. 

The final shiny app can be found in this [repo](https://github.com/aj2duncan/leaflet-draw){target="_blank"} and clicking on the [Binder](https://mybinder.org/){target="_blank"} badge will launch things. 




