---
title: Extracting Columns from a Tibble
author: ''
date: '2020-02-18'
slug: extracting-columns-from-a-tibble
categories:
  - R
tags:
  - tibble
description: ''
highlight: true
---


```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, comment = NA, 
                      warning = FALSE, message = FALSE)
```

This is both a follow up post to my earlier post [tibbles, data frames and vectors](https://www.aj2duncan.com/blog/tibbles-dataframes-and-vectors/) and a query that came up at work. 

The central question was how can you extract a column from a tibble as part of a pipe. That might sound straight forward but let's take a look.

```{r}
# load packages and create a tibble to work with
library(tibble)
library(dplyr)
test = tibble(id = c("a", "b", "a"), val = c(1, 2, 3))
test
```

Let's run a quick `filter()` and then `select()` the values.

```{r}
test %>%
  filter(id == "a") %>%
  select(val)
```

But as we saw in my previous post, this is still a tibble, so if we pass it to something expecting a vector it doesn't work.

```{r, warning = TRUE}
test %>%
  filter(id == "a") %>%
  select(val) %>%
  mean()
```

## The Best Solution

The best solution to this problem is to use the `pull()` function. I missed this function when it first came out in [dplyr 0.7.0](https://blog.rstudio.com/2017/06/13/dplyr-0-7-0/) but now use it a lot.

```{r}
# pull() gives us a vector 
test %>%
  filter(id == "a") %>%
  pull(val) 

# now we don't get an error
test %>%
  filter(id == "a") %>%
  pull(val) %>%
  mean()
```

## Other Solutions

As with many, many things in `R` there are alternatives and for the moment we are just going to continue to work with a pipe. Some neat alternatives come from a special type of function called an [infix](http://adv-r.had.co.nz/Functions.html#special-calls) function. You use these all the time, `+` and `%>%` being examples. These *infix* functions sit between arguments `x + y` but they can also be used in a more traditional `f(x,y)` manner. To use them in this way you need to use backticks so `2^3 = 8` becomes `` `^`(2,3) = 8`` :

```{r}
# Power
2 ^ 3
`^`(2,3)
```

This means that we can use `[[]]` and `$` like this too. Take our tibble.

```{r}
test$val
`$`(test, val)
```

or 

```{r}
test[[2]]
`[[`(test, 2)
```

This means we can add an infix function to the end of our pipe.

```{r}
test %>%
  filter(id == "a") %>%
  `[[`(2)
```

```{r}
test %>%
  filter(id == "a") %>%
  `$`(val)
```

Finally, we could just use `getElement()`.

```{r}
test %>%
  filter(id == "a") %>%
  getElement("val")
```

