---
title: Using `dplyr::rename()` within a function
author: Andrew Duncan
date: '2020-06-10'
slug: using-dplyr-rename-within-a-function
categories:
  - R
tags:
  - tibble
  - dplyr
description: ''
highlight: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, comment = NA, 
                      warning = FALSE, message = FALSE)
```


Today I rewrote some old code and came across something that I thought I would write down because I'm pretty sure I'll forget it if I don't! The issue was passing strings to the `rename()` function from the [dplyr](https://dplyr.tidyverse.org/){target='_blank} package to rename a column. This came about because of a change in a older^[much older] version of dplyr which had functions specifically for use inside your own functions. These had an underscore as a postfix. So instead of using `rename()` I was using `rename_()`. There are now depreciated and so I was updating my code.

It was a pretty simple function but still created a few errors until I found the necessary bit in the documentation. More on that in a minute, first let's create some data and see what we're trying to do.

```{r}
library(dplyr)
temp <- tribble(
  ~x, ~y,
  1, 2,
  3, 4
)
temp
```


Now we've got some data we can `rename()` a column.

```{r}
rename(temp, x2 = x)
```

So far so good. However, things get a little trickier if you want to do this inside another function. For example we could try this function.

```{r}
my_rename <- function(dat_f, new_name) {
  rename(dat_f, new_name = x)
}
```

We can try using this with a string and without.

```{r}
my_rename(temp, "x2")
my_rename(temp, x2)
```

Not quite what we were looking for and it doesn't produce an error. After a little looking I found the necessary information on the use of *tidy selection* in `rename()` - you can read it [here](https://dplyr.tidyverse.org/articles/programming.html){target='_blank'}. With the new info we can rewrite the function to use [glue](https://glue.tidyverse.org){target='_blank'} notation and assignment with the `:=` operator. This works with and without quotes.

```{r}
my_rename <- function(dat_f, new_name) {
  rename(dat_f, {{new_name}} := x)
}
my_rename(temp, "x2")
my_rename(temp, x3)
```

In a future post I plan to do a run through (mainly for myself when I forget things) a few things like this including *quosures*.